%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}{\slshape
    Ognun vede quel che tu pari. Pochi sentono quel che tu sei.
        \\
        --- Niccolò Machiavelli}
\end{flushright}



\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}

Many people contributed to this work and to my studies and I would like to
thank them here.

\bigskip

First of all I would like to thank the Politecnico of Milano and its
professors. In particular I would like to thank the NECST Lab, for the many
corrections, for the (malriposta) trust and opportunities, and for the freedom
it gave me in pursuing my interests and goals.


\bigskip

CERN, my colleagues and my supervisors that for some weird reasons seem to
appreciate all the mess I am leaving in their codebase. Accelerating!

\bigskip

The friends from China whose support and friendship helped me in these years
and that were always there to cazzeggiare. Sapervi cosi’ distanti ma cosi’
uniti non mi fa temere nessuna distanza, grazie!

\bigskip

The new friends from Milan who have never let me down. Parlare con
voi mi ha fatto crescere, grazie!

\bigskip

The old friends of a lifetime that always make me laugh, talk, share and that
are always there to sfottermi. Siete sempre in grado di ricordarmi chi sono, da
dove vengo e sopratutto dove vado, grazie!

\bigskip

Then to my family, that always managed to catch me when I was trying
to run away as a child, but then exhausted eventually let me go in order: 
to the US, to Milan, to China and last to Switzerland. Per
l’incodizionato supporto in ogni mia scelta e decisione, grazie!

\bigskip

My brother, I don't know exactly why, but Thanks. Mi fai vedere il mondo da un
altra prospettiva senza mai farmi sentire solo, grazie!

\bigskip

Finally Ema (Miao Miao) who in these years has always been there for me and with me,
her presence and smile make me stronger and bolder allowing me to pursue
goals that I don’t know if I would have been able to reach without her. Ti mao!


\endgroup
