%************************************************
\chapter{Conclusion and Future Work}\label{ch:FutureWorks}
%************************************************

In this work we have presented a way to merge efficient run time dependency
management using containers technologies with efficient content distribution
provided by CVMFS. 

The proposed methodology exposed on Chapter \ref{ch:Methodology} is based on
providing a POSIX file system that container runtime can use to load their
content. The use of containers allow to efficiently manage runtime dependencies
logically packing them in different layers exposed as directories.  We build
such file system on CVMFS to exploit its efficient content distribution
mechanism. This forced some of our design decision regarding the file system
tree like the use of \textit{super directories}, however the exact same
structure is applicable also to different file system, not only to CVMFS. 

The implementation of the proposed methodology have been explored on Chapter
\ref{ch:Implementation} where we have described a software capable of creating
and managing the whole proposed file system from ingesting the containers
content to deleting it without breaking old images. Moreover we have concluded
the work started with the \textit{cvmfs/graphdriver} that provided only a way
to use the \textit{thin images} while we finally proposed a methodology to
easily produce them starting from standard Docker \textit{fat} image and deploy
the \textit{thin images} on standard Docker Registries for an easy
distribution.

Finally, on Chapter \ref{ch:Results} we have shown the advantages that the
proposed system brings in terms of startup time of uncached containers and
bandwidth consumption while at the same time not imposing run time penalties.

The final work is in a stable state and there are plans to actually deploy it
in production inside CERN, nevertheless several further advancement are
possible with respect to increasing the set of supported container run-times or
improving the software that manages the file system.

For what concerns the run-times we believe that similar work to the one done
for Docker with the \textit{cvmfs/graphdriver} plugin can be done also for
different technologies. In particular we are confident that the approach of
\textit{thin images} can be adopted also by \textit{containerd}
\cite{containerd}, this time implementing a custom \textit{snapshotter} instead
of a Docker graphdriver plugin. Different container run-times needs further
investigation.

A big limitation to lazily serve the content for containers using lazy systems
like CVMFS is the OCI standard. Container run-times that follow the OCI
standard do not expose any interface to actually load custom content into the
container itself, but they rely simply on tarballs to provide the content,
hence \textit{hacks} like the one in the \textit{cvmfs/graphdriver} are
necessary. A solution could be a custom container run-time which provides the
possibility to specify how the content should be loaded and from where.

Another improvement to the system is related to the tool that builds the file
system structure. It could be re-structured to limit the cyclomatic complexity
in order to allow further enhancement. Another big improvement would be to
exploit to the fullest the transactional interface of CVMFS. Indeed several
transactions are used when ingesting a new image into the file system. An
improvement would be to use only a single transaction per image.

